---
title: "Dplyr"
author: "Jennifer Brosnahan"
date: "7/31/2020"
output: 
  html_document: 
    keep_md: yes
---
## More dplyr practice because we can

### Dplyr uses the following functions:
#+ filter() and slice()
#+ arrange()
#+ select() and rename()
#+ distinct()
#+ mutate() and transmute()
#+ summarise()
#+ sample_n() and sample_frac()

```{r, warning=FALSE}
library(dplyr)
library(nycflights13)
```

```{r}
head(flights)
```

```{r}
summary(flights)
```

## Filter(df, column1 arg, column2 arg, ...) much easier syntax
```{r}
head(filter(flights, month==11, day==3, carrier=='AA'))
```


## Slice function: used to slice out rows by position
```{r}
slice(flights, 10:20)
```

## Arrange, instead of filtering or selecting rows, it reorders row
```{r}
head(arrange(flights, year, month, day, arr_time))
```

## Arrange in descending order by passing desc function to whatever column you want descending
```{r}
head(arrange(flights, year, month, day, desc(arr_time)))
```

## Select function: Quickly grab data and columns you want to select
```{r}
head(select(flights, carrier))
```
```{r}
head(select(flights, carrier, arr_time))
```

## Rename function: allows you to quickly rename columns
```{r}
head(rename(flights, airline_carrier = carrier))
```

## Distinct function: allows you to select unique values in column or table
```{r}
distinct(select(flights, carrier))
```

### As you can see, distinct provided the 16 unique carrier values in df

## Mutate function: Add new columns that are built off a formula(s) of existing columns
```{r}
head(mutate(flights, new_col = arr_delay - dep_delay))
```

### If you only want the new_col back, use transmute() instead of mutate()

## Summarise function: allows you to quickly collapse df into single rows using function that aggregates result
```{r}
summarise(flights, avg_air_time = mean(air_time, na.rm = TRUE))
```

### na.rm = TRUE is for mean function to just remove na values

```{r}
summarise(flights, total_time = sum(air_time, na.rm = TRUE)) 
```

### Total time spent in air by all flights

## Sample_n, sample_frac: allows you to take random number of rows or fraction of rows
```{r}
sample_n(flights, 10) # 10 random rows
```

```{r}
sample_frac(flights, .005) # 10% of random rows
```

## Pipe operator is totally cool!

### Difference between usual nesting syntax vs. pipe. Nesting is more difficult read (from inside out):
```{r}
df <- mtcars
```

```{r}
result <- arrange(sample_n(filter(df, mpg>20), size=5), desc(mpg))
result
```


### Pipe formula: df %>% op1 %>% op2 %>% op3
```{r}
df %>% filter(mpg>20) %>% 
  sample_n(5) %>% 
  arrange(desc(mpg))
```

## Dplyr training exercises using mtcars
```{r}
head(mtcars)
```

### Return rows of cars that have an mpg value greater than 20 and 6 cylinders
```{r}
head(filter(mtcars, mpg>20, cyl==6))
```

### Reorder df by cyl first, then by descending wt.
```{r}
arrange(mtcars, cyl, desc(wt))
```

### Select the columns mpg and hp
```{r}
select(mtcars, mpg, hp)
```

### Select the distinct values of the gear column
```{r}
distinct(mtcars, gear)
```

### Create a new column called "Performance" which is calculated by hp divided by wt
```{r}
mutate(mtcars, Performance = hp / wt)
```

### Find the mean mpg value using dplyr
```{r}
summarise(mtcars, avg_mpg = mean(mpg))
```

### Use pipe operators to get the mean hp value for cars with 6 cylinders
```{r}
mtcars %>% filter(cyl==6) %>% 
  summarise(mean(hp))
```
